---
title: "Dense scenario issue - univariate models"
author: "Fabio Morgante"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

```{r set opts, message=FALSE}
library(mr.mash.alpha)
library(glmnet)
library(mr.ash.alpha)

###Set options
options(stringsAsFactors = FALSE)

###Set seed
RNGversion("3.5.0")
set.seed(123)
```

Let's simulate data with n=600, p=1,000, p_causal=500, r=5, r_causal=5, PVE=0.5, shared effects, independent predictors, and independent residuals. The models will be fitted to the full data.

```{r simulate data}
###Set parameters
n <- 600
p <- 1000
p_causal <- 500
r <- 5
r_causal <- r
pve <- 0.5
B_cor <- 1
X_cor <- 0
V_cor <- 0

###Simulate V, B, X and Y
out <- mr.mash.alpha:::simulate_mr_mash_data(n, p, p_causal, r, r_causal=r, intercepts = rep(1, r),
                                             pve=pve, B_cor=B_cor, B_scale=0.9, X_cor=X_cor, X_scale=0.8, V_cor=V_cor)

colnames(out$Y) <- paste0("Y", seq(1, r))
rownames(out$Y) <- paste0("N", seq(1, n))
colnames(out$X) <- paste0("X", seq(1, p))
rownames(out$X) <- paste0("N", seq(1, n))
```

We run *glmnet* with $\alpha=1$ to obtain an inital estimate for the regression coefficients to provide to *mr.ash*. Then, we run *mr.mash* initialized from the *mr.ash* solution. All the methods are run with standardize=FALSE.

```{r fit, results='hide'}
###Compute univariate summary stats needed to compute the prior variances
univ_sumstats <- mr.mash.alpha:::get_univariate_sumstats(out$X, out$Y, standardize=FALSE, standardize.response=FALSE)

###Define matrices to store the coefficients
Bhat_glmnet_univ <- matrix(as.numeric(NA), nrow=p, ncol=r)
Bhat_mrash <- matrix(as.numeric(NA), nrow=p, ncol=r)
Bhat_mrmash_univ <- matrix(as.numeric(NA), nrow=p, ncol=r)

###Loop through responses
for(resp in 1:r){
  ###Fit lasso
  cvfit_glmnet_univ <- cv.glmnet(x=out$X, y=out$Y[, resp], family="gaussian", alpha=1, standardize=FALSE)
  coeff_glmnet_univ <- coef(cvfit_glmnet_univ, s="lambda.min")
  Bhat_glmnet_univ[, resp] <- as.vector(coeff_glmnet_univ)[-1]
  
  ###Compute prior variances
  grid_univ <- c(0, mr.mash.alpha:::autoselect.mixsd(univ_sumstats$Bhat[, resp], univ_sumstats$Shat[, resp], mult=2))^2
  
  ###Fit mr.ash
  fit_mrash <- mr.ash.alpha::mr.ash(out$X, out$Y[, resp], sa2=grid_univ, standardize=FALSE, update.pi=TRUE, update.sigma=TRUE,
                                    beta.init=Bhat_glmnet_univ[, resp])
  Bhat_mrash[, resp] <- drop(fit_mrash$beta)
  
  ###Fit mr.mash
  S0 <- vector("list", length(grid_univ))
  for(i in 1:length(grid_univ)){
    S0[[i]] <- matrix(grid_univ[i], ncol=1, nrow=1) 
  }

  fit_mrmash_univ <- mr.mash(out$X, matrix(out$Y[, resp], ncol=1), S0, tol=1e-2, convergence_criterion="ELBO", update_w0=TRUE, 
                            update_w0_method="EM", standardize=FALSE, verbose=FALSE, update_V=TRUE, update_V_method="full", 
                            w0_threshold=0, mu1_init=fit_mrash$beta)
  Bhat_mrmash_univ[, resp] <- drop(fit_mrmash_univ$mu1)
}
```

Let's look at the results for each response.

```{r plot results, fig.height=12, fig.width=12}
###Response 1
resp <- 1

layout(matrix(c(1, 1, 2, 2,
                1, 1, 2, 2,
                0, 3, 3, 0,
                0, 3, 3, 0), 4, 4, byrow = TRUE))

###Plot estimated vs true coeffcients
ymax <- max(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
ymin <- min(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
##mr.mash
plot(out$B[, resp], Bhat_mrmash_univ[, resp], main="mr.mash", xlab="True coefficients", ylab="Estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, ylim=c(ymin, ymax))
##mr.ash
plot(out$B[, resp], Bhat_mrash[, resp], main="mr.ash", xlab="True coefficients", ylab="Estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, ylim=c(ymin, ymax))

###Plot mr.mash vs glmnet estimated coeffcients
colorz <- matrix("black", nrow=p, ncol=1)
zeros <- out$B[, resp]==0
for(i in 1:ncol(colorz)){
  colorz[zeros, i] <- "red"
}

xymax <- max(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
xymin <- min(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))

plot(Bhat_mrash[, resp], Bhat_mrmash_univ[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     col=colorz, cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)
legend("topleft", 
       legend = c("Non-zero", "Zero"), 
       col = c("black", "red"), 
       pch = c(1, 1), 
       horiz = FALSE,
       cex=2)

###Response 2
resp <- 2

layout(matrix(c(1, 1, 2, 2,
                1, 1, 2, 2,
                0, 3, 3, 0,
                0, 3, 3, 0), 4, 4, byrow = TRUE))

###Plot estimated vs true coeffcients
ymax <- max(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
ymin <- min(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
##mr.mash
plot(out$B[, resp], Bhat_mrmash_univ[, resp], main="mr.mash", xlab="True coefficients", ylab="Estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, ylim=c(ymin, ymax))
##mr.ash
plot(out$B[, resp], Bhat_mrash[, resp], main="mr.ash", xlab="True coefficients", ylab="Estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, ylim=c(ymin, ymax))

###Plot mr.mash vs glmnet estimated coeffcients
colorz <- matrix("black", nrow=p, ncol=1)
zeros <- out$B[, resp]==0
for(i in 1:ncol(colorz)){
  colorz[zeros, i] <- "red"
}

xymax <- max(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
xymin <- min(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))

plot(Bhat_mrash[, resp], Bhat_mrmash_univ[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     col=colorz, cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)
legend("topleft", 
       legend = c("Non-zero", "Zero"), 
       col = c("black", "red"), 
       pch = c(1, 1), 
       horiz = FALSE,
       cex=2)

###Response 3
resp <- 3

layout(matrix(c(1, 1, 2, 2,
                1, 1, 2, 2,
                0, 3, 3, 0,
                0, 3, 3, 0), 4, 4, byrow = TRUE))

###Plot estimated vs true coeffcients
ymax <- max(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
ymin <- min(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
##mr.mash
plot(out$B[, resp], Bhat_mrmash_univ[, resp], main="mr.mash", xlab="True coefficients", ylab="Estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, ylim=c(ymin, ymax))
##mr.ash
plot(out$B[, resp], Bhat_mrash[, resp], main="mr.ash", xlab="True coefficients", ylab="Estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, ylim=c(ymin, ymax))

###Plot mr.mash vs glmnet estimated coeffcients
colorz <- matrix("black", nrow=p, ncol=1)
zeros <- out$B[, resp]==0
for(i in 1:ncol(colorz)){
  colorz[zeros, i] <- "red"
}

xymax <- max(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
xymin <- min(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))

plot(Bhat_mrash[, resp], Bhat_mrmash_univ[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     col=colorz, cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)
legend("topleft", 
       legend = c("Non-zero", "Zero"), 
       col = c("black", "red"), 
       pch = c(1, 1), 
       horiz = FALSE,
       cex=2)

###Response 4
resp <- 4

layout(matrix(c(1, 1, 2, 2,
                1, 1, 2, 2,
                0, 3, 3, 0,
                0, 3, 3, 0), 4, 4, byrow = TRUE))

###Plot estimated vs true coeffcients
ymax <- max(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
ymin <- min(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
##mr.mash
plot(out$B[, resp], Bhat_mrmash_univ[, resp], main="mr.mash", xlab="True coefficients", ylab="Estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, ylim=c(ymin, ymax))
##mr.ash
plot(out$B[, resp], Bhat_mrash[, resp], main="mr.ash", xlab="True coefficients", ylab="Estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, ylim=c(ymin, ymax))

###Plot mr.mash vs glmnet estimated coeffcients
colorz <- matrix("black", nrow=p, ncol=1)
zeros <- out$B[, resp]==0
for(i in 1:ncol(colorz)){
  colorz[zeros, i] <- "red"
}

xymax <- max(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
xymin <- min(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))

plot(Bhat_mrash[, resp], Bhat_mrmash_univ[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     col=colorz, cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)
legend("topleft", 
       legend = c("Non-zero", "Zero"), 
       col = c("black", "red"), 
       pch = c(1, 1), 
       horiz = FALSE,
       cex=2)

###Response 5
resp <- 5

layout(matrix(c(1, 1, 2, 2,
                1, 1, 2, 2,
                0, 3, 3, 0,
                0, 3, 3, 0), 4, 4, byrow = TRUE))

###Plot estimated vs true coeffcients
ymax <- max(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
ymin <- min(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
##mr.mash
plot(out$B[, resp], Bhat_mrmash_univ[, resp], main="mr.mash", xlab="True coefficients", ylab="Estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, ylim=c(ymin, ymax))
##mr.ash
plot(out$B[, resp], Bhat_mrash[, resp], main="mr.ash", xlab="True coefficients", ylab="Estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, ylim=c(ymin, ymax))

###Plot mr.mash vs glmnet estimated coeffcients
colorz <- matrix("black", nrow=p, ncol=1)
zeros <- out$B[, resp]==0
for(i in 1:ncol(colorz)){
  colorz[zeros, i] <- "red"
}

xymax <- max(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))
xymin <- min(c(Bhat_mrmash_univ[, resp], Bhat_mrash[, resp]))

plot(Bhat_mrash[, resp], Bhat_mrmash_univ[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     col=colorz, cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)
legend("topleft", 
       legend = c("Non-zero", "Zero"), 
       col = c("black", "red"), 
       pch = c(1, 1), 
       horiz = FALSE,
       cex=2)
```

As we can see, *mr.mash* shrinks large coeffcients more than *mr.ash*. Now, we need to understand why these two implementations are giving different results. We will start by making sure that the two methods give the same results with fixed residual variance (and properly scaled prior variances in the case of *mr.ash*).

```{r fit fixed residual variance, results='hide'}
###Define matrices to store the coefficients
Bhat_mrash_fixV <- matrix(as.numeric(NA), nrow=p, ncol=r)
Bhat_mrmash_univ_fixV <- matrix(as.numeric(NA), nrow=p, ncol=r)

###Loop through responses
for(resp in 1:r){
  ###Compute prior variances
  grid_univ <- c(0, mr.mash.alpha:::autoselect.mixsd(univ_sumstats$Bhat[, resp], univ_sumstats$Shat[, resp], mult=2))^2
  
  ###Fit mr.ash
  fit_mrash_fixV <-mr.ash(out$X, out$Y[, resp], standardize=FALSE, update.pi=TRUE, update.sigma=FALSE, sa2 = grid_univ/var(out$Y[,resp]),
                          beta.init=Bhat_glmnet_univ[, resp], sigma2 = var(out$Y[,resp]), pi = rep(1/length(grid_univ), length(grid_univ)))
  Bhat_mrash_fixV[, resp] <- drop(fit_mrash_fixV$beta)
  
  ###Fit mr.mash
  S0 <- vector("list", length(grid_univ))
  for(i in 1:length(grid_univ)){
    S0[[i]] <- matrix(grid_univ[i], ncol=1, nrow=1) 
  }

  fit_mrmash_univ_fixV <- mr.mash(out$X, matrix(out$Y[, resp], ncol=1), S0, tol=1e-4, convergence_criterion="ELBO", update_w0=TRUE, 
                            update_w0_method="EM", standardize=FALSE, verbose=FALSE, update_V=FALSE, update_V_method="full", 
                            w0_threshold=0, mu1_init=fit_mrash_fixV$beta, V=matrix(var(out$Y[,resp]), 1, 1), e=0)
  Bhat_mrmash_univ_fixV[, resp] <- drop(fit_mrmash_univ_fixV$mu1)
}
```

Let's look at the results for each response.

```{r plot results fixed V, fig.height=12, fig.width=12}
layout(matrix(c(1, 1, 2, 2, 3, 3,
                1, 1, 2, 2, 3, 3,
                0, 4, 4, 5, 5, 0,
                0, 4, 4, 5, 5, 0), 4, 6, byrow = TRUE))

resp <- 1

xymax <- max(c(Bhat_mrmash_univ_fixV[, resp], Bhat_mrash_fixV[, resp]))
xymin <- min(c(Bhat_mrmash_univ_fixV[, resp], Bhat_mrash_fixV[, resp]))

plot(Bhat_mrash_fixV[, resp], Bhat_mrmash_univ_fixV[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)

resp <- 2

xymax <- max(c(Bhat_mrmash_univ_fixV[, resp], Bhat_mrash_fixV[, resp]))
xymin <- min(c(Bhat_mrmash_univ_fixV[, resp], Bhat_mrash_fixV[, resp]))

plot(Bhat_mrash_fixV[, resp], Bhat_mrmash_univ_fixV[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)

resp <- 3

xymax <- max(c(Bhat_mrmash_univ_fixV[, resp], Bhat_mrash_fixV[, resp]))
xymin <- min(c(Bhat_mrmash_univ_fixV[, resp], Bhat_mrash_fixV[, resp]))

plot(Bhat_mrash_fixV[, resp], Bhat_mrmash_univ_fixV[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)

resp <- 4

xymax <- max(c(Bhat_mrmash_univ_fixV[, resp], Bhat_mrash_fixV[, resp]))
xymin <- min(c(Bhat_mrmash_univ_fixV[, resp], Bhat_mrash_fixV[, resp]))

plot(Bhat_mrash_fixV[, resp], Bhat_mrmash_univ_fixV[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)

resp <- 5

xymax <- max(c(Bhat_mrmash_univ_fixV[, resp], Bhat_mrash_fixV[, resp]))
xymin <- min(c(Bhat_mrmash_univ_fixV[, resp], Bhat_mrash_fixV[, resp]))

plot(Bhat_mrash_fixV[, resp], Bhat_mrmash_univ_fixV[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)
```

The results look very similar, so the coordinate ascent algorithm and the update of the mixture weights should not be the culprit of the differences seen in the first plots. N.B. we still had to make the convergence criterion stricter in *mr.mash* to reach this level of agreement.

After a conversation, Matthew and I discovered that I have been providing a grid that was not sensible to *mr.ash*. This is because of the different (scaled) parameterization of the prior, $b \sim N(0, \sigma^2 \sigma^2_b)$. This might account for all the differences between *mr.ash* and *mr.mash* we have seen. To test this hypothesis, we will use this strategy:
1. Fit *mr.ash* with a sufficiently dense and wide grid.
2. Fit *mr.mash* with the same grid multiplied by the estimated residual variance from *mr.ash*
In this way, the two grids will be approximately equivalent, given the two different parameterizations of the model.

```{r fit equivalent grid, results='hide'}
###Compute grid of prior variances for mr.ash
grid_mrash <- c(0, sort(10^seq(2,-3.5,length.out = 500)))

###Define matrices to store the coefficients
Bhat_mrash_eqgrid <- matrix(as.numeric(NA), nrow=p, ncol=r)
Bhat_mrmash_univ_eqgrid <- matrix(as.numeric(NA), nrow=p, ncol=r)

###Loop through responses
for(resp in 1:r){

  ###Fit mr.ash
  fit_mrash_eqgrid <- mr.ash.alpha::mr.ash(out$X, out$Y[, resp], sa2=grid_mrash, standardize=FALSE, update.pi=TRUE, update.sigma=TRUE,
                                    beta.init=Bhat_glmnet_univ[, resp])
  Bhat_mrash_eqgrid[, resp] <- drop(fit_mrash_eqgrid$beta)
  
  ###Fit mr.mash
  S0 <- list()
  for(i in 1:length(grid_mrash)){
    S0[[i]] <- matrix(grid_mrash[i]*fit_mrash_eqgrid$sigma2, ncol=1, nrow=1) 
  }
  
  fit_mrmash_univ_eqgrid <- mr.mash(out$X, matrix(out$Y[, resp], ncol=1), S0, tol=1e-4, convergence_criterion="ELBO", update_w0=TRUE, 
                                    update_w0_method="EM", standardize=FALSE, verbose=FALSE, update_V=TRUE, update_V_method="full", 
                                    w0_threshold=0, mu1_init=matrix(Bhat_glmnet_univ[, resp], ncol=1), e=0)  
  Bhat_mrmash_univ_eqgrid[, resp] <- drop(fit_mrmash_univ_eqgrid$mu1)
}
```

```{r plot results equivalent grid, fig.height=12, fig.width=12}
layout(matrix(c(1, 1, 2, 2, 3, 3,
                1, 1, 2, 2, 3, 3,
                0, 4, 4, 5, 5, 0,
                0, 4, 4, 5, 5, 0), 4, 6, byrow = TRUE))

resp <- 1

xymax <- max(c(Bhat_mrmash_univ_eqgrid[, resp], Bhat_mrash_eqgrid[, resp]))
xymin <- min(c(Bhat_mrmash_univ_eqgrid[, resp], Bhat_mrash_eqgrid[, resp]))

plot(Bhat_mrash_eqgrid[, resp], Bhat_mrmash_univ_eqgrid[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)

resp <- 2

xymax <- max(c(Bhat_mrmash_univ_eqgrid[, resp], Bhat_mrash_eqgrid[, resp]))
xymin <- min(c(Bhat_mrmash_univ_eqgrid[, resp], Bhat_mrash_eqgrid[, resp]))

plot(Bhat_mrash_eqgrid[, resp], Bhat_mrmash_univ_eqgrid[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)

resp <- 3

xymax <- max(c(Bhat_mrmash_univ_eqgrid[, resp], Bhat_mrash_eqgrid[, resp]))
xymin <- min(c(Bhat_mrmash_univ_eqgrid[, resp], Bhat_mrash_eqgrid[, resp]))

plot(Bhat_mrash_eqgrid[, resp], Bhat_mrmash_univ_eqgrid[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)

resp <- 4

xymax <- max(c(Bhat_mrmash_univ_eqgrid[, resp], Bhat_mrash_eqgrid[, resp]))
xymin <- min(c(Bhat_mrmash_univ_eqgrid[, resp], Bhat_mrash_eqgrid[, resp]))

plot(Bhat_mrash_eqgrid[, resp], Bhat_mrmash_univ_eqgrid[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)

resp <- 5

xymax <- max(c(Bhat_mrmash_univ_eqgrid[, resp], Bhat_mrash_eqgrid[, resp]))
xymin <- min(c(Bhat_mrmash_univ_eqgrid[, resp], Bhat_mrash_eqgrid[, resp]))

plot(Bhat_mrash_eqgrid[, resp], Bhat_mrmash_univ_eqgrid[, resp], main="mr.mash vs mr.ash", 
     xlab="mr.ash estimated coefficients", ylab="mr.mash estimated coefficients",
     cex=2, cex.lab=1.8, cex.main=2, cex.axis=1.8, xlim=c(xymin, xymax), ylim=c(xymin, xymax))
abline(0, 1)
```

The results are now essentially the same! The remaining differences (I believe) can be attributed to the slightly different solutions achieved due to the different convergence checks used in the two software (i.e., ELBO in `mr.mash` and mixture weights in `mr.ash`).

So, the take-home messages are: 
1. At least in this example, the two different parameterizations lead essentially to the same results.
2. One needs to be careful in choosing the grid of prior variances, **especially** with the scaled parameterization, which is less intuitive!