---
title: "mr.mash vs mtlasso vs UTMOST"
author: "Fabio Morgante"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---

```{r set opts, message=FALSE}
###Set options
options(stringsAsFactors=FALSE)

###Load libraries
library(dscrutils)
library(cowplot)
library(ggplot2)
library(scales)

###Function to convert dscquery output from list to data.frame suitable for plotting
convert_dsc_to_dataframe <- function(dsc){
  ###Data.frame to store the results after convertion
  dsc_df <- data.frame()

  ###Get length of list elements
  n_elem <- length(dsc$DSC)

  ###Loop through the dsc list
  for(i in 1:n_elem){
    ##Prepare vectors making up the final data frame
    r_scalar <- dsc$simulate.r[i]
    repp <- rep(dsc$DSC[i], times=r_scalar)
    n <- rep(dsc$simulate.n[i], times=r_scalar)
    p <- rep(dsc$simulate.p[i], times=r_scalar)
    p_causal <- rep(dsc$simulate.p_causal[i], times=r_scalar)
    r <- rep(dsc$simulate.r[i], times=r_scalar)
    response <- 1:r_scalar
    pve <- rep(dsc$simulate.pve[i], times=r_scalar)
    simulate <- rep(dsc$simulate[i], times=r_scalar)
    fit <- rep(dsc$fit[i], times=r_scalar)
    score <- rep(dsc$score[i], times=r_scalar)
    score.err <- dsc$score.err[[i]]
    timing <- rep(dsc$fit.time[i], times=r_scalar)

    ##Build the data frame
    df <- data.frame(rep=repp, n=n, p=p, p_num_caus=p_causal, r=r, response=response, pve=pve,
                     scenario=simulate, method=fit, score_metric=score, score_value=score.err, time=timing)
    dsc_df <- rbind(dsc_df, df)
  }

  return(dsc_df)
}

###Function to compute rmse (relative to mr_mash_consec_em)
compute_rrmse <- function(dsc_plot, log10_scale=FALSE){
  dsc_plot <- transform(dsc_plot, experiment=paste(rep, response, scenario, sep="-"))
  t <- 0
  for (i in unique(dsc_plot$experiment)) {
    t <- t+1
    rmse_data  <- dsc_plot[which(dsc_plot$experiment == i & dsc_plot$score_metric=="scaled_rmse"), ]
    mse_mr_mash_consec_em <- rmse_data[which(rmse_data$method=="mr_mash"), "score_value"]
    if(!log10_scale)
      rmse_data$score_value <- rmse_data$score_value/mse_mr_mash_consec_em
    else
      rmse_data$score_value <- log10(rmse_data$score_value/mse_mr_mash_consec_em)
    rmse_data$score_metric <- "scaled_rrmse"
    if(t>1){
      rmse_data_tot <- rbind(rmse_data_tot, rmse_data)
    } else if(t==1){
      rmse_data_tot <- rmse_data
    }
  }
  
  rmse_data_tot$experiment <- NULL
  
  return(rmse_data_tot)
}

colors <- c("skyblue", "limegreen", "red")
```

In all the analyses so far, we have used [Abhishek's implementation of the sparse multi-task LASSO](https://users.rcc.uchicago.edu/~aksarkar/gtex-pred/lasso.html) (called mtlasso here), the method that was used in [UTMOST](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6788740/). This is because the [original implementation](https://github.com/yiminghu/CTIMP) requires the data in an awkward format. Note that Abhishek's implementation is not exactly the same as UTMOST. For example, the initialization of the coefficients and the way to choose the grid of values for the tuning parameters are different. Recently, the Ritchie Lab reimplemented the [method](https://github.com/RitchieLab/multi_tissue_twas_sim) using the original code from the UTMOST autors so that it takes in data in the standard format (i.e., Y in matrix form). The caveat is that this reimplementation does not allow for missing values in Y. In this website, we want to compare *mr.mash* to both the Ritchie Lab's implementation and Abhishek's implementation of UTMOST. 

We simulated five scenarios with n=500, p=1000, r=10, PVE=0.2, p_causal=randomly sampled between 1 and 10, independent predictors, independent residuals, and the following structure of  effects across conditions:

* equal effects.

* independent effects.

* effects only in condition 1.

* equal effects only in condition 1-3.

* shared effects in condition 1-3, shared effects in condition 7-10 (with per-condition PVE=0.05), and no sharing between these two groups. 

*mr.mash* was fitted with only canonical covariance matrices, updating V (imposing a diagonal structure), and updating the prior weights using EM updates. Because UTMOST only centers the predictors, mtlasso and *mr.mash* were also run only centering the predictors for consistency. 

The accuracy of prediction is measured as the root mean squared error (RMSE) scaled by the standard deviation of Y.

### Equal effects across conditions

```{r equal effects simulations, message=FALSE, fig.height=12, fig.width=15}
dsc_out <- dscquery("output/mvreg_mrmash_vs_mtlasso_vs_utmost", 
		    c("simulate.n", "simulate.p", "simulate.p_causal", "simulate.r",   
                      "simulate.pve", "simulate.B_cor", "simulate.B_scale",
                      "simulate.X_cor", "simulate.X_scale",
                      "simulate.V_cor", "simulate.prop_testset",
                      "simulate", "fit", "score", "score.err", "fit.time"), 
                    conditions = "$(simulate) == 'indepX_indepV_sharedB_allr_norm'", verbose=FALSE)

dsc_plots <- convert_dsc_to_dataframe(dsc_out)
rrmse_dat <- compute_rrmse(dsc_plots)
dsc_plots <- rbind(dsc_plots, rrmse_dat)

dsc_plots$method_fac <- factor(dsc_plots$method, levels=c("mr_mash", "mtlasso", "utmost"))

p <- ggplot(dsc_plots[which(dsc_plots$score_metric=="scaled_rrmse"), ], aes_string(x = "method_fac", y = "score_value", fill = "method_fac")) +
  	geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
  	scale_fill_manual(values = colors) +
  	labs(x = "", y = "Scaled RRMSE", fill="Method") +
  	theme_cowplot(font_size = 20) +
  	theme(axis.line.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank())

print(p)

###Create plot for run time
dsc_plots_time <- dsc_plots[which(dsc_plots$response==1 & dsc_plots$score_metric=="scaled_rrmse"),
                            -which(colnames(dsc_plots) %in% c("score_metric", "score_value", "response"))]

p_time <- ggplot(dsc_plots_time, aes_string(x = "method", y = "time", fill = "method")) +
  geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(trans="log2", breaks = trans_breaks("log2", function(x) 2^x),
                     labels = trans_format("log2", math_format(2^.x))) +
  labs(x = "", y = "Elapsed time (seconds) in log2 scale", title = "Run time", fill="Method") +
  theme_cowplot(font_size = 20) +
  theme(axis.line.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))

print(p_time)
```

### Independent effects across conditions

```{r independent effects simulations, message=FALSE, fig.height=12, fig.width=15}
dsc_out <- dscquery("output/mvreg_mrmash_vs_mtlasso_vs_utmost",
                    c("simulate.n", "simulate.p", "simulate.p_causal", "simulate.r",
                      "simulate.pve", "simulate.B_cor", "simulate.B_scale",
                      "simulate.X_cor", "simulate.X_scale",
                      "simulate.V_cor", "simulate.prop_testset",
                      "simulate", "fit", "score", "score.err", "fit.time"),
                    conditions = "$(simulate) == 'indepX_indepV_indepB_allr_norm'", verbose=FALSE)

dsc_plots <- convert_dsc_to_dataframe(dsc_out)
rrmse_dat <- compute_rrmse(dsc_plots)
dsc_plots <- rbind(dsc_plots, rrmse_dat)

dsc_plots$method_fac <- factor(dsc_plots$method, levels=c("mr_mash", "mtlasso", "utmost"))

p <- ggplot(dsc_plots[which(dsc_plots$score_metric=="scaled_rrmse"), ], aes_string(x = "method_fac", y = "score_value", fill = "method_fac")) +
        geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
        scale_fill_manual(values = colors) +
        labs(x = "", y = "Scaled RRMSE", fill="Method") +
        theme_cowplot(font_size = 20) +
        theme(axis.line.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank())

print(p)

###Create plot for run time
dsc_plots_time <- dsc_plots[which(dsc_plots$response==1 & dsc_plots$score_metric=="scaled_rrmse"),
                            -which(colnames(dsc_plots) %in% c("score_metric", "score_value", "response"))]

p_time <- ggplot(dsc_plots_time, aes_string(x = "method", y = "time", fill = "method")) +
  geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(trans="log2", breaks = trans_breaks("log2", function(x) 2^x),
                     labels = trans_format("log2", math_format(2^.x))) +
  labs(x = "", y = "Elapsed time (seconds) in log2 scale", title = "Run time", fill="Method") +
  theme_cowplot(font_size = 20) +
  theme(axis.line.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))

print(p_time)
```

### Effects only in condition 1

```{r singleton effects simulations, message=FALSE, fig.height=12, fig.width=15}
dsc_out <- dscquery("output/mvreg_mrmash_vs_mtlasso_vs_utmost",
                    c("simulate.n", "simulate.p", "simulate.p_causal", "simulate.r",
                      "simulate.pve", "simulate.B_cor", "simulate.B_scale",
                      "simulate.X_cor", "simulate.X_scale",
                      "simulate.V_cor", "simulate.prop_testset",
                      "simulate", "fit", "score", "score.err", "fit.time"),
                    conditions = "$(simulate) == 'indepX_indepV_B_1causalr'", verbose=FALSE)

dsc_plots <- convert_dsc_to_dataframe(dsc_out)
rrmse_dat <- compute_rrmse(dsc_plots)
dsc_plots <- rbind(dsc_plots, rrmse_dat)

dsc_plots$method_fac <- factor(dsc_plots$method, levels=c("mr_mash", "mtlasso", "utmost"))
dsc_plots$response_fac <- as.factor(dsc_plots$response)

p <- ggplot(dsc_plots[which(dsc_plots$score_metric=="scaled_rrmse"), ], aes_string(x = "response_fac", y = "score_value", fill = "method_fac")) +
        geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
        scale_fill_manual(values = colors) +
        labs(x = "Condition", y = "Scaled RRMSE", fill="Method") +
        theme_cowplot(font_size = 20)

print(p)

###Create plot for run time
dsc_plots_time <- dsc_plots[which(dsc_plots$response==1 & dsc_plots$score_metric=="scaled_rrmse"),
                            -which(colnames(dsc_plots) %in% c("score_metric", "score_value", "response"))]

p_time <- ggplot(dsc_plots_time, aes_string(x = "method", y = "time", fill = "method")) +
  geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(trans="log2", breaks = trans_breaks("log2", function(x) 2^x),
                     labels = trans_format("log2", math_format(2^.x))) +
  labs(x = "", y = "Elapsed time (seconds) in log2 scale", title = "Run time", fill="Method") +
  theme_cowplot(font_size = 20) +
  theme(axis.line.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))

print(p_time)
```

### Equal effects in condition 1-3

```{r one block effects simulations, message=FALSE, fig.height=12, fig.width=15}
dsc_out <- dscquery("output/mvreg_mrmash_vs_mtlasso_vs_utmost",
                    c("simulate.n", "simulate.p", "simulate.p_causal", "simulate.r",
                      "simulate.pve", "simulate.B_cor", "simulate.B_scale",
                      "simulate.X_cor", "simulate.X_scale",
                      "simulate.V_cor", "simulate.prop_testset",
                      "simulate", "fit", "score", "score.err", "fit.time"),
                    conditions = "$(simulate) == 'indepX_indepV_sharedB_3causalrespr10'", verbose=FALSE)

dsc_plots <- convert_dsc_to_dataframe(dsc_out)
rrmse_dat <- compute_rrmse(dsc_plots)
dsc_plots <- rbind(dsc_plots, rrmse_dat)

dsc_plots$method_fac <- factor(dsc_plots$method, levels=c("mr_mash", "mtlasso", "utmost"))
dsc_plots$response_fac <- as.factor(dsc_plots$response)

p <- ggplot(dsc_plots[which(dsc_plots$score_metric=="scaled_rrmse"), ], aes_string(x = "response_fac", y = "score_value", fill = "method_fac")) +
        geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
        scale_fill_manual(values = colors) +
        labs(x = "Condition", y = "Scaled RRMSE", fill="Method") +
        theme_cowplot(font_size = 20)

print(p)

###Create plot for run time
dsc_plots_time <- dsc_plots[which(dsc_plots$response==1 & dsc_plots$score_metric=="scaled_rrmse"),
                            -which(colnames(dsc_plots) %in% c("score_metric", "score_value", "response"))]

p_time <- ggplot(dsc_plots_time, aes_string(x = "method", y = "time", fill = "method")) +
  geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(trans="log2", breaks = trans_breaks("log2", function(x) 2^x),
                     labels = trans_format("log2", math_format(2^.x))) +
  labs(x = "", y = "Elapsed time (seconds) in log2 scale", title = "Run time", fill="Method") +
  theme_cowplot(font_size = 20) +
  theme(axis.line.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))

print(p_time)
```

### Shared effects in condition 1-3, shared effects in condition 7-10 (with per-condition PVE=0.05), and no sharing between these two groups

```{r two blocks effects simulations, message=FALSE, fig.height=12, fig.width=15}
dsc_out <- dscquery("output/mvreg_mrmash_vs_mtlasso_vs_utmost",
                    c("simulate.n", "simulate.p", "simulate.p_causal", "simulate.r",
                      "simulate.pve", "simulate.B_cor", "simulate.B_scale",
                      "simulate.X_cor", "simulate.X_scale",
                      "simulate.V_cor", "simulate.prop_testset",
                      "simulate", "fit", "score", "score.err", "fit.time"),
                    conditions = "$(simulate) == 'indepX_indepV_sharedB_2blocksr_diffPVE'", verbose=FALSE)

dsc_plots <- convert_dsc_to_dataframe(dsc_out)
rrmse_dat <- compute_rrmse(dsc_plots)
dsc_plots <- rbind(dsc_plots, rrmse_dat)

dsc_plots$method_fac <- factor(dsc_plots$method, levels=c("mr_mash", "mtlasso", "utmost"))
dsc_plots$response_fac <- as.factor(dsc_plots$response)

p <- ggplot(dsc_plots[which(dsc_plots$score_metric=="scaled_rrmse"), ], aes_string(x = "response_fac", y = "score_value", fill = "method_fac")) +
        geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
        scale_fill_manual(values = colors) +
        labs(x = "Condition", y = "Scaled RRMSE", fill="Method") +
        theme_cowplot(font_size = 20)

print(p)

###Create plot for run time
dsc_plots_time <- dsc_plots[which(dsc_plots$response==1 & dsc_plots$score_metric=="scaled_rrmse"),
                            -which(colnames(dsc_plots) %in% c("score_metric", "score_value", "response"))]

p_time <- ggplot(dsc_plots_time, aes_string(x = "method", y = "time", fill = "method")) +
  geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(trans="log2", breaks = trans_breaks("log2", function(x) 2^x),
                     labels = trans_format("log2", math_format(2^.x))) +
  labs(x = "", y = "Elapsed time (seconds) in log2 scale", title = "Run time", fill="Method") +
  theme_cowplot(font_size = 20) +
  theme(axis.line.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))

print(p_time)
```
The results show that *mr.mash* estimates the coefficients better than both implementations of the sparse multi-task LASSO and is much faster. Abhishek's implementation produces results that agree with UTMOST in most scenatios, but seems slower (especially since the grid for the tuning parameters has length = 10 while it has length = 100 in UTMOST). That makes sense since the coordinate updates are implemented in C++ in UTMOST. 
