---
title: "Convergence"
author: "Fabio Morgante"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

```{r set opts}
options(stringsAsFactors = FALSE)
```

##Simulation 1 -- Shared effects, independent variables

```{r load data 1}
dat1 <- readRDS("output/fit_mr_mash_n600_p1000_p_caus50_r5_pve0.5_sigmaoffdiag1_sigmascale0.8_gammaoffdiag0_gammascale0.8_Voffdiag0.2_Vscale0_updatew0TRUE_updatew0TRUE_updatew0methodmixsqp_updateVTRUE.rds")
n1 <- dat1$params$n
p1 <- dat1$params$p
p_causal1 <- dat1$params$p_causal
r1 <- dat1$params$r
k1 <- length(dat1$fit$w0)
pve1 <- dat1$params$pve
prop_testset1 <- dat1$params$prop_testset
progress_dat1 <- dat1$fit$progress
V1 <- dat1$inputs$V
Sigma1 <- dat1$inputs$Sigma
Gamma1 <- dat1$inputs$Gamma
```

The results below are based on simulation with `r n1` samples, `r p1` variables of which `r p_causal1` were causal, `r r1` responses with a per-response proportion of variance explained (PVE) of `r pve1`. Variables, X, were drawn from MVN(0, Gamma), causal effects, B, were drawn from MVN(0, Sigma). The responses, Y, were drawn from MN(XB, I, V).

```{r disp corrs 1}
cat("Gamma (First 5 elements)")
Gamma1[1:5, 1:5]

cat("Sigma")
Sigma1

cat("V")
V1
```

mr.mash was fitted to the training data (`r (1-prop_testset1)*100`% of the data) updating V and updating the prior weights using mixSQP. The mixture prior consisted of `r k1` components.

Here, we investigate convergence. Convergence was reached when max($mu1_{t}$ - $mu1_{t-1}$) was less than 1e-8.

```{r ELBO 1, fig.height=12, fig.width=15}
plot(progress_dat1$iter, progress_dat1$ELBO_diff, xlab="Iteration", ylab="log Difference in ELBO", main="ELBO vs iteration", type="b", pch=16, cex.lab=1.5, cex.axis=1.5, log="y")
```

```{r mu1 1, fig.height=12, fig.width=15}
plot(progress_dat1$iter, progress_dat1$mu1_max.diff, xlab="Iteration", ylab="log max(Difference in mu1)", main="mu1 vs iteration", type="b", pch=16, cex.lab=1.5, cex.axis=1.5, log="y")
```


##Simulation 2 -- Independent effects, independent variables

```{r load data 2}
dat2 <- readRDS("output/fit_mr_mash_n600_p1000_p_caus50_r5_pve0.5_sigmaoffdiag0_sigmascale0.8_gammaoffdiag0_gammascale0.8_Voffdiag0.2_Vscale0_updatew0TRUE_updatew0TRUE_updatew0methodmixsqp_updateVTRUE.rds")
n2 <- dat2$params$n
p2 <- dat2$params$p
p_causal2 <- dat2$params$p_causal
r2 <- dat2$params$r
k2 <- length(dat2$fit$w0)
pve2 <- dat2$params$pve
prop_testset2 <- dat2$params$prop_testset
progress_dat2 <- dat2$fit$progress
V2 <- dat2$inputs$V
Sigma2 <- dat2$inputs$Sigma
Gamma2 <- dat2$inputs$Gamma
```

The results below are based on simulation with `r n2` samples, `r p2` variables of which `r p_causal2` were causal, `r r2` responses with a per-response proportion of variance explained (PVE) of `r pve2`. Variables, X, were drawn from MVN(0, Gamma), causal effects, B, were drawn from MVN(0, Sigma). The responses, Y, were drawn from MN(XB, I, V).

```{r disp corrs 2}
cat("Gamma (First 5 elements)")
Gamma2[1:5, 1:5]

cat("Sigma")
Sigma2

cat("V")
V2
```

mr.mash was fitted to the training data (`r (1-prop_testset2)*100`% of the data) updating V and updating the prior weights using mixSQP. The mixture prior consisted of `r k2` components.

Here, we investigate convergence. Convergence was reached when max($mu1_{t}$ - $mu1_{t-1}$) was less than 1e-8.

```{r ELBO 2, fig.height=12, fig.width=15}
plot(progress_dat2$iter, progress_dat2$ELBO_diff, xlab="Iteration", ylab="log Difference in ELBO", main="ELBO vs iteration", type="b", pch=16, cex.lab=1.5, cex.axis=1.5, log="y")
```

```{r mu1 2, fig.height=12, fig.width=15}
plot(progress_dat2$iter, progress_dat2$mu1_max.diff, xlab="Iteration", ylab="log max(Difference in mu1)", main="mu1 vs iteration", type="b", pch=16, cex.lab=1.5, cex.axis=1.5, log="y")
```


##Simulation 3 -- Shared effects, correlated variables

```{r load data 3}
dat3 <- readRDS("output/fit_mr_mash_n600_p1000_p_caus50_r5_pve0.5_sigmaoffdiag1_sigmascale0.8_gammaoffdiag0.5_gammascale0.8_Voffdiag0.2_Vscale0_updatew0TRUE_updatew0TRUE_updatew0methodmixsqp_updateVTRUE.rds")
n3 <- dat3$params$n
p3 <- dat3$params$p
p_causal3 <- dat3$params$p_causal
r3 <- dat3$params$r
k3 <- length(dat3$fit$w0)
pve3 <- dat3$params$pve
prop_testset3 <- dat3$params$prop_testset
progress_dat3 <- dat3$fit$progress
V3 <- dat3$inputs$V
Sigma3 <- dat3$inputs$Sigma
Gamma3 <- dat3$inputs$Gamma
```

The results below are based on simulation with `r n3` samples, `r p3` variables of which `r p_causal3` were causal, `r r3` responses with a per-response proportion of variance explained (PVE) of `r pve3`. Variables, X, were drawn from MVN(0, Gamma), causal effects, B, were drawn from MVN(0, Sigma). The responses, Y, were drawn from MN(XB, I, V).

```{r disp corrs 3}
cat("Gamma (First 5 elements)")
Gamma3[1:5, 1:5]

cat("Sigma")
Sigma3

cat("V")
V3
```

mr.mash was fitted to the training data (`r (1-prop_testset3)*100`% of the data) updating V and updating the prior weights using mixSQP. The mixture prior consisted of `r k3` components.

Here, we investigate convergence. Convergence was reached when max($mu1_{t}$ - $mu1_{t-1}$) was less than 1e-8.

```{r ELBO 3, fig.height=12, fig.width=15}
plot(progress_dat3$iter, progress_dat3$ELBO_diff, xlab="Iteration", ylab="log Difference in ELBO", main="ELBO vs iteration", type="b", pch=16, cex.lab=1.5, cex.axis=1.5, log="y")
```

```{r mu1 3, fig.height=12, fig.width=15}
plot(progress_dat3$iter, progress_dat3$mu1_max.diff, xlab="Iteration", ylab="log max(Difference in mu1)", main="mu1 vs iteration", type="b", pch=16, cex.lab=1.5, cex.axis=1.5, log="y")
```


##Simulation 4 -- Independent effects, correlated variables

```{r load data 4}
dat4 <- readRDS("output/fit_mr_mash_n600_p1000_p_caus50_r5_pve0.5_sigmaoffdiag0_sigmascale0.8_gammaoffdiag0.5_gammascale0.8_Voffdiag0.2_Vscale0_updatew0TRUE_updatew0TRUE_updatew0methodmixsqp_updateVTRUE.rds")
n4 <- dat4$params$n
p4 <- dat4$params$p
p_causal4 <- dat4$params$p_causal
r4 <- dat4$params$r
k4 <- length(dat4$fit$w0)
pve4 <- dat4$params$pve
prop_testset4 <- dat4$params$prop_testset
progress_dat4 <- dat4$fit$progress
V4 <- dat4$inputs$V
Sigma4 <- dat4$inputs$Sigma
Gamma4 <- dat4$inputs$Gamma
```

The results below are based on simulation with `r n4` samples, `r p4` variables of which `r p_causal4` were causal, `r r4` responses with a per-response proportion of variance explained (PVE) of `r pve4`. Variables, X, were drawn from MVN(0, Gamma), causal effects, B, were drawn from MVN(0, Sigma). The responses, Y, were drawn from MN(XB, I, V).

```{r disp corrs 4}
cat("Gamma (First 5 elements)")
Gamma4[1:5, 1:5]

cat("Sigma")
Sigma4

cat("V")
V4
```

mr.mash was fitted to the training data (`r (1-prop_testset4)*100`% of the data) updating V and updating the prior weights using mixSQP. The mixture prior consisted of `r k4` components.

Here, we investigate convergence. Convergence was reached when max($mu1_{t}$ - $mu1_{t-1}$) was less than 1e-8.

```{r ELBO 4, fig.height=12, fig.width=15}
plot(progress_dat4$iter, progress_dat4$ELBO_diff, xlab="Iteration", ylab="log Difference in ELBO", main="ELBO vs iteration", type="b", pch=16, cex.lab=1.5, cex.axis=1.5, log="y")
```

```{r mu1 4, fig.height=12, fig.width=15}
plot(progress_dat4$iter, progress_dat4$mu1_max.diff, xlab="Iteration", ylab="log max(Difference in mu1)", main="mu1 vs iteration", type="b", pch=16, cex.lab=1.5, cex.axis=1.5, log="y")
```
