---
title: "Effects structure with 3 causal responses out of 10"
author: "Fabio Morgante"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---

```{r set opts, warning=FALSE, message=FALSE}
###Set options
options(stringsAsFactors=FALSE)

###Load libraries
library(dscrutils)
library(ggplot2)
library(cowplot)
library(scales)

###Function to convert dscquery output from list to data.frame suitable for plotting
convert_dsc_to_dataframe <- function(dsc){
  ###Data.frame to store the results after convertion
  dsc_df <- data.frame()
  
  ###Get length of list elements 
  n_elem <- length(dsc$DSC)
  
  ###Loop through the dsc list
  for(i in 1:n_elem){
    ##Prepare vectors making up the final data frame
    r_scalar <- dsc$simulate.r[i]
    repp <- rep(dsc$DSC[i], times=r_scalar)
    n <- rep(dsc$simulate.n[i], times=r_scalar)
    p <- rep(dsc$simulate.p[i], times=r_scalar)
    p_causal <- rep(dsc$simulate.p_causal[i], times=r_scalar)
    r <- rep(dsc$simulate.r[i], times=r_scalar)
    response <- 1:r_scalar
    pve <- rep(dsc$simulate.pve[i], times=r_scalar)
    simulate <- rep(dsc$simulate[i], times=r_scalar)
    fit <- rep(dsc$fit[i], times=r_scalar)
    score <- rep(dsc$score[i], times=r_scalar)
    score.err <- dsc$score.err[[i]]
    timing <- rep(dsc$fit.time[i], times=r_scalar)
    
    ##Build the data frame
    df <- data.frame(rep=repp, n=n, p=p, p_num_caus=p_causal, r=r, response=response, pve=pve,  
                     scenario=simulate, method=fit, score_metric=score, score_value=score.err, time=timing)
    dsc_df <- rbind(dsc_df, df)
  }
  
  return(dsc_df)
}

###Function to compute rmse (relative to mr_mash_consec_em)
compute_rrmse <- function(dsc_plot, log10_scale=FALSE){
  dsc_plot <- transform(dsc_plot, experiment=paste(rep, response, scenario, sep="-"))
  t <- 0
  for (i in unique(dsc_plot$experiment)) {
    t <- t+1
    rmse_data  <- dsc_plot[which(dsc_plot$experiment == i & dsc_plot$score_metric=="scaled_mse"), ]
    mse_mr_mash_consec_em <- rmse_data[which(rmse_data$method=="mr_mash_em_can"), "score_value"]
    if(!log10_scale)
      rmse_data$score_value <- rmse_data$score_value/mse_mr_mash_consec_em
    else
      rmse_data$score_value <- log10(rmse_data$score_value/mse_mr_mash_consec_em)
    rmse_data$score_metric <- "rrmse"
    if(t>1){
      rmse_data_tot <- rbind(rmse_data_tot, rmse_data)
    } else if(t==1){
      rmse_data_tot <- rmse_data
    }
  }
  
  rmse_data_tot$experiment <- NULL
  
  return(rmse_data_tot)
}

###Function to shift legend in the empty facet
shift_legend <- function(p) {
  library(gtable)
  library(lemon)
  # check if p is a valid object
  if(!(inherits(p, "gtable"))){
    if(inherits(p, "ggplot")){
      gp <- ggplotGrob(p) # convert to grob
    } else {
      message("This is neither a ggplot object nor a grob generated from ggplotGrob. Returning original plot.")
      return(p)
    }
  } else {
    gp <- p
  }
  
  # check for unfilled facet panels
  facet.panels <- grep("^panel", gp[["layout"]][["name"]])
  empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]]), 
                               USE.NAMES = F)
  empty.facet.panels <- facet.panels[empty.facet.panels]
  
  if(length(empty.facet.panels) == 0){
    message("There are no unfilled facet panels to shift legend into. Returning original plot.")
    return(p)
  }
  
  # establish name of empty panels
  empty.facet.panels <- gp[["layout"]][empty.facet.panels, ]
  names <- empty.facet.panels$name
  
  # return repositioned legend
  reposition_legend(p, 'center', panel=names)
}

###Set some quantities used in the following plots
colors <- c("skyblue", "dodgerblue", "limegreen", "green", "gold", "orange", "red", "firebrick", "darkmagenta", "mediumpurple")
facet_labels <- c(r2 = "r2", bias = "bias", rrmse="rRMSE")
```

```{r load data indepX, warning=FALSE, message=FALSE}
###Load the dsc results
dsc_out <- dscquery("output/mvreg_all_genes_prior_indepX_indepV_sharedB_3causalrespr10", 
                    c("simulate.n", "simulate.p", "simulate.p_causal", "simulate.r",   
                      "simulate.w","simulate.r_causal", "simulate.pve", "simulate.B_cor",  
                      "simulate.B_scale", "simulate.X_cor", "simulate.X_scale",
                      "simulate.V_cor", "simulate", "fit", "score", "score.err", "fit.time"), 
                    groups="fit: mr_mash_em_can, mr_mash_em_data, mr_mash_em_dataAndcan, mlasso, mridge, menet", 
                    verbose=FALSE)

###Obtain simulation parameters
prop_testset <- 0.2
n <- unique(dsc_out$simulate.n)
p <- unique(dsc_out$simulate.p)
p_causal <- unique(dsc_out$simulate.p_causal)
r <- unique(dsc_out$simulate.r)
r_causal <- eval(parse(text=unique(dsc_out$simulate.r_causal)))
pve <- unique(dsc_out$simulate.pve)
w <- eval(parse(text=unique(dsc_out$simulate.w)))
B_cor <- eval(parse(text=unique(dsc_out$simulate.B_cor)))
B_scale <- eval(parse(text=unique(dsc_out$simulate.B_scale)))
X_cor <- unique(dsc_out$simulate.X_cor)
X_scale <- unique(dsc_out$simulate.X_scale)
V_cor <- unique(dsc_out$simulate.V_cor)

###Remove list elements that are not useful anymore
dsc_out$simulate.r_causal <- NULL
dsc_out$simulate.w <- NULL
dsc_out$simulate.B_cor <- NULL
dsc_out$simulate.B_scale <- NULL
dsc_out$simulate.X_cor <- NULL
dsc_out$simulate.X_scale <- NULL
dsc_out$simulate.V_cor <- NULL
```

# Independent predictors

## Simulation set up

The results below are based on simulations with `r n` samples, `r p` variables of which `r p_causal` were causal, `r r` responses with a per-response proportion of variance explained (PVE) of `r pve`. Variables, X, were drawn from MVN(0, Gamma), where Gamma is such that it achieves a correlation between variables of `r X_cor` and a scale of `r X_scale`. Causal effects, B, were drawn from MVN(0, Sigma), Sigma is such that it achieves a correlation between responses of `r B_cor[1]`. Response `r r_causal[[1]]` have causal effects while the remaining seven responses do not have any causal effect. The responses, Y, were drawn from MN(XB, I, V), where V is such that it achieves a correlation between responses of `r V_cor` and a scale defined by PVE, in the case of cuasal responses. Non causal responses are from N(0, 1).

2000 such datasets (i.e., "genes") were simulated and univariate summary statistics were obtained by simple linear regression in the training data (`r (1-prop_testset)*100`% of the data. To mirror what would happen in real data analysis, the indexes of the training-test individuals were the same for all the datasets. However, since these 2000 datasets were simulated independently, I do not think it matters). These regression coefficients and standard errors were used as input in the *mash* pipeline (from Gao) to compute data-driven covariance matrices (up to the ED step included). In particular, the top variable per dataset was used to define a "strong" set and 4 random variables per dataset were used to define a "random" set. Covariance matrices were estimated using *flash*, PCA (including the top 3 PCs), and the empirical covariance matrix.  

The first 50 datasets were used for the prediction analysis. *mr.mash* was fitted to the training data, updating V (imposing a diagonal structure) and updating the prior weights using EM updates. The mixture prior consisted of components defined by:

* canonical matrices correpsonding to different settings of effect sharing/specificity (i.e., singletons, independent, low heterogeneity, medium heterogeneity, high heterogeneity, shared) plus the spike. 
* data-driven matrices estimated as described above plus the spike.
* both canonical and data-driven matrices plus the spike.

The covariance matrices were scaled by a grid of values computed from the univariate summary statistics as in the *mash* paper. The posterior mean of the regression coefficients were initialized to the estimates of the group-LASSO. The mixture weights were initialized with the proportion of zero-coefficients from the group-LASSO estimate as the weight on the spike and the proportion of non-zero-coefficients split equally among the remaining components.Convergence was declared when the maximum difference in the ELBO between two successive iterations was smaller than 1e-2. 

Then, responses were predicted on the test data (`r prop_testset*100`% of the data). 

Here, we evaluate the accuracy of prediction assessed by $r^2$ and bias (slope) from the regression of the true response on the predicted response, and the relative root mean square error (rRMSE) scaled by the standard deviation of the true responses in the test data. The boxplots are across the 50 datasets, for each response separately.

## mr.mash vs other methods

Here, we compare *mr.mash* to the multivariate versions of LASSO as implemented in *glmnet*. The form of the penalty id the following: $\lambda[(1-\alpha)/2 ||\mathbf{\beta}_j||^2_2 + \alpha ||\mathbf{\beta}_j||_2]$. $\lambda$ is chosen by cross-validation in the training set. All the methods were using 4 threads -- *mr.mash* loops over the mixture components in parallel, *glmnet* loops over folds in parallel.

```{r accuracy vs others indepX, fig.height=12, fig.width=15, message=FALSE}
###Convert from list to data.frame for plotting
dsc_plots <- convert_dsc_to_dataframe(dsc_out)

###Compute rmse score (relative to mr_mash_consec_em) and add it to the data
rrmse_dat <- compute_rrmse(dsc_plots)
dsc_plots <- rbind(dsc_plots, rrmse_dat)

###Remove mse from scores and keep only methods wanted
dsc_plots <- dsc_plots[which(dsc_plots$score_metric!="scaled_mse" ), ]
dsc_plots <- dsc_plots[which(dsc_plots$method %in% c("mr_mash_em_can", "mr_mash_em_data",
                                                          "mr_mash_em_dataAndcan", "mlasso")), ]

###Create factor version of method
dsc_plots$method_fac <- factor(dsc_plots$method, levels=c("mr_mash_em_can", "mr_mash_em_data",
                                                          "mr_mash_em_dataAndcan", "mlasso"),
                                                labels=c("mr_mash_can", "mr_mash_data", "mr_mash_both",
                                                         "mlasso"))

###Build data.frame with best accuracy achievable
hlines <- data.frame(score_metric=c("r2", "bias", "rrmse"), max_val=c(unique(dsc_plots$pve), 1, 1))

###Create factor version of response
dsc_plots$response_fac <- as.factor(dsc_plots$response)

###Plot bias
p_bias <- ggplot(dsc_plots[which(dsc_plots$score_metric=="bias"), ], aes_string(x = "response_fac", y = "score_value", fill = "method_fac")) +
  geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
  scale_fill_manual(values = colors) +
  labs(x = "Response", y = "Bias", title = "Bias", fill="Method") +
  geom_hline(data=hlines[which(hlines$score_metric=="bias"), ], aes(yintercept = max_val), linetype="dashed", size=1) +
  theme_cowplot(font_size = 20) +
  theme(plot.title = element_text(hjust = 0.5))

print(p_bias)

###Plot r2
p_r2 <- ggplot(dsc_plots[which(dsc_plots$score_metric=="r2"), ], aes_string(x = "response_fac", y = "score_value", fill = "method_fac")) +
  geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
  scale_fill_manual(values = colors) +
  labs(x = "Response", y = "r2", title = "r2", fill="Method") +
  geom_hline(data=hlines[which(hlines$score_metric=="r2"), ], aes(yintercept = max_val), linetype="dashed", size=1) +
  theme_cowplot(font_size = 20) +
  theme(plot.title = element_text(hjust = 0.5))

print(p_r2)

###Plot rrmse
p_rrmse <- ggplot(dsc_plots[which(dsc_plots$score_metric=="rrmse"), ], aes_string(x = "response_fac", y = "score_value", fill = "method_fac")) +
  geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
  scale_fill_manual(values = colors) +
  labs(x = "Response", y = "rRMSE", title = "rMSE", fill="Method") +
  geom_hline(data=hlines[which(hlines$score_metric=="rrmse"), ], aes(yintercept = max_val), linetype="dashed", size=1) +
  theme_cowplot(font_size = 20) +
  theme(plot.title = element_text(hjust = 0.5))

print(p_rrmse)
```

Let's now remove outliers from the plots to make things a little clearer.

```{r accuracy no outliers indepX, fig.height=12, fig.width=15, message=FALSE}
###Plot bias
p_bias <- ggplot(dsc_plots[which(dsc_plots$score_metric=="bias"), ], aes_string(x = "response_fac", y = "score_value", fill = "method_fac")) +
  Ipaper::geom_boxplot2(color = "black",width.errorbar = 0, width = 0.85) +  
  scale_fill_manual(values = colors) +
  labs(x = "Response", y = "Bias", title = "Bias", fill="Method") +
  geom_hline(data=hlines[which(hlines$score_metric=="bias"), ], aes(yintercept = max_val), linetype="dashed", size=1) +
  theme_cowplot(font_size = 20) +
  theme(plot.title = element_text(hjust = 0.5))

print(p_bias)

###Plot r2
p_r2 <- ggplot(dsc_plots[which(dsc_plots$score_metric=="r2"), ], aes_string(x = "response_fac", y = "score_value", fill = "method_fac")) +
  Ipaper::geom_boxplot2(color = "black",width.errorbar = 0, width = 0.85) +  
  scale_fill_manual(values = colors) +
  labs(x = "Response", y = "r2", title = "r2", fill="Method") +
  geom_hline(data=hlines[which(hlines$score_metric=="r2"), ], aes(yintercept = max_val), linetype="dashed", size=1) +
  theme_cowplot(font_size = 20) +
  theme(plot.title = element_text(hjust = 0.5))

print(p_r2)

###Plot rrmse
p_rrmse <- ggplot(dsc_plots[which(dsc_plots$score_metric=="rrmse"), ], aes_string(x = "response_fac", y = "score_value", fill = "method_fac")) +
  Ipaper::geom_boxplot2(color = "black",width.errorbar = 0, width = 0.85) +
  scale_fill_manual(values = colors) +
  labs(x = "Response", y = "rRMSE", title = "rMSE", fill="Method") +
  geom_hline(data=hlines[which(hlines$score_metric=="rrmse"), ], aes(yintercept = max_val), linetype="dashed", size=1) +
  theme_cowplot(font_size = 20) +
  theme(plot.title = element_text(hjust = 0.5))

print(p_rrmse)
```

Here, we look at the elapsed time ($log_{10}$ seconds) of each method. Note that the *mr.mash* run time does not include the run time of group-LASSO (but should be considered since we used it to initialize *mr.mash*).

```{r run time indepX, fig.height=12, fig.width=15, message=FALSE}
dsc_plots_time <- dsc_plots[which(dsc_plots$response==1 & dsc_plots$score_metric=="r2"), 
                          -which(colnames(dsc_plots) %in% c("score_metric", "score_value", "response"))]

p_time <- ggplot(dsc_plots_time, aes_string(x = "method_fac", y = "time", fill = "method_fac")) +
  geom_boxplot(color = "black", outlier.size = 1, width = 0.85) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(trans="log10", breaks = trans_breaks("log10", function(x) 10^x),
                      labels = trans_format("log10", math_format(10^.x))) +
  labs(x = "", y = "Elapsed time (seconds) in log10 scale",title = "Run time", fill="Method") +
  theme_cowplot(font_size = 20) +
  theme(axis.line.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))

print(p_time)
```



In summary, *mr.mash* with the data-driven covariance matrices does very well and runs pretty fast. Using only the canonical covariance matrices is not as effective (as expected) in this case, but it's still good. Using both types of covariance matrices does not add anything in terms of performance to using only the data-driven matrices. However, it makes the method much slower to run.