---
title: "Analysis of 10 GTEx genes"
author: "Fabio Morgante"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---

```{r set opts, message=FALSE}
###Set options
options(stringsAsFactors=FALSE)

###Load libraries
library(cowplot)
library(ggplot2)

###Function to compute accuracy
compute_accuracy <- function(Y, Yhat) {
  bias <- rep(as.numeric(NA), ncol(Y))
  names(bias) <- colnames(Y)
  r2 <- rep(as.numeric(NA), ncol(Y))
  names(r2) <- colnames(Y)
  scaled_rmse <- rep(as.numeric(NA), ncol(Y))
  names(scaled_rmse) <- colnames(Y)
  
  for(i in 1:ncol(Y)){ 
    dat <- na.omit(data.frame(Y[, i], Yhat[, i]))
    colnames(dat) <- c("Y", "Yhat")
    
    fit  <- lm(Y ~ Yhat, data=dat)
    bias[i] <- coef(fit)[2] 
    r2[i] <- summary(fit)$r.squared
    scaled_rmse[i] <- sqrt(mean((dat$Y - dat$Yhat)^2))/sd(dat$Y)
  }
  
  return(list(bias=bias, r2=r2, scaled_rmse=scaled_rmse))
}

###Genes to be analyzed
genes <- c("ENSG00000083812.11", "ENSG00000156413.13", "ENSG00000171119.2",
           "ENSG00000198131.13", "ENSG00000214626.2", "ENSG00000268201.1",
	   "ENSG00000268866.1", "ENSG00000269794.1", "ENSG00000269855.2",
	   "ENSG00000273901.1")

###Set some parameters
thresh <- 100
```

Here, we compare the prediction performance of *mr.mash*, [Abhishek's implementation of the sparse multi-task LASSO](https://github.com/aksarkar/mtlasso), and the univariate elastic net. The analysis was performed on 10 randomly sampled GTEx v8 genes using the following pipelines: 

* [univariate summary statistics computation](https://github.com/cumc/bioworkflows/blob/master/multivariate-prediction/cv_preprocessing.ipynb)
* [data-driven covariance matrices computation](https://github.com/cumc/bioworkflows/blob/master/multivariate-fine-mapping/mixture_prior.ipynb)
* [prior grid computation](https://github.com/cumc/bioworkflows/blob/master/multivariate-prediction/prior_grid.ipynb)
* [*mr.mash* model fitting](https://github.com/cumc/bioworkflows/blob/master/multivariate-prediction/mrmash.ipynb)
* [sparse multi-task LASSO model fitting](https://github.com/cumc/bioworkflows/blob/master/multivariate-prediction/mtlasso.ipynb) 

Briefly, we used a 5-fold cross validation scheme, whereby all the steps but the computation of prediction performance were done on the training set. Only data-driven covariance matrices were used and computed using *flash* (unconstrained and non-negative factors only) and PCA, denoised using the TED method in the [udr R package](https://github.com/stephenslab/udr). A common grid of scaling factors for all the genes was computed from their univariate summary statistics. These two steps used all the GTEx genes expressed in all the 49 tissues. Regression coefficients and prior weights were initialized using the univariate elastic net. Prior weights were updated in a gene specific manner, as was the residual covariance V with no diagonal constrain. We also tested a "second pass" version of *mr.mash* where we combined the responsibilities from the "first pass" across genes and performed a joint weights update. The jointly updated weights were fixed for all the genes. Prediction performance was evaluated as the Root Mean Squared Error (RMSE) scaled by the variance of true expression in the test set. The plots below show the mean RMSE across folds for each tissue for pairs of methods.     

```{r process results}
p_mrmash_sec <- vector("list", length(genes))
p_enet <- vector("list", length(genes))
p_mtlasso <- vector("list", length(genes))

t <- 0

for(gene in genes){

    t <- t+1

    ###Load the data
    dat1_first <- readRDS(paste0("output/gtex_mr_mash_analysis/prediction/fold_1/", gene, ".GTEx_V8_fold_1_batch_1_mrmash.first_pass.rds"))
    dat2_first <- readRDS(paste0("output/gtex_mr_mash_analysis/prediction/fold_2/", gene, ".GTEx_V8_fold_2_batch_1_mrmash.first_pass.rds"))
    dat3_first <- readRDS(paste0("output/gtex_mr_mash_analysis/prediction/fold_3/", gene, ".GTEx_V8_fold_3_batch_1_mrmash.first_pass.rds"))
    dat4_first <- readRDS(paste0("output/gtex_mr_mash_analysis/prediction/fold_4/", gene, ".GTEx_V8_fold_4_batch_1_mrmash.first_pass.rds"))
    dat5_first <- readRDS(paste0("output/gtex_mr_mash_analysis/prediction/fold_5/", gene, ".GTEx_V8_fold_5_batch_1_mrmash.first_pass.rds"))

    dat1_sec <- readRDS(paste0("output/gtex_mr_mash_analysis/prediction/fold_1/", gene, ".GTEx_V8_fold_1_batch_1_mrmash.second_pass.rds"))
    dat2_sec <- readRDS(paste0("output/gtex_mr_mash_analysis/prediction/fold_2/", gene, ".GTEx_V8_fold_2_batch_1_mrmash.second_pass.rds"))
    dat3_sec <- readRDS(paste0("output/gtex_mr_mash_analysis/prediction/fold_3/", gene, ".GTEx_V8_fold_3_batch_1_mrmash.second_pass.rds"))
    dat4_sec <- readRDS(paste0("output/gtex_mr_mash_analysis/prediction/fold_4/", gene, ".GTEx_V8_fold_4_batch_1_mrmash.second_pass.rds"))
    dat5_sec <- readRDS(paste0("output/gtex_mr_mash_analysis/prediction/fold_5/", gene, ".GTEx_V8_fold_5_batch_1_mrmash.second_pass.rds"))

    dat1_mtlasso <- readRDS(paste0("output/gtex_mr_mash_analysis/mtlasso/fold_1/", gene, ".GTEx_V8_fold_1_mtlasso.rds"))
    dat2_mtlasso <- readRDS(paste0("output/gtex_mr_mash_analysis/mtlasso/fold_2/", gene, ".GTEx_V8_fold_2_mtlasso.rds"))
    dat3_mtlasso <- readRDS(paste0("output/gtex_mr_mash_analysis/mtlasso/fold_3/", gene, ".GTEx_V8_fold_3_mtlasso.rds"))
    dat4_mtlasso <- readRDS(paste0("output/gtex_mr_mash_analysis/mtlasso/fold_4/", gene, ".GTEx_V8_fold_4_mtlasso.rds"))
    dat5_mtlasso <- readRDS(paste0("output/gtex_mr_mash_analysis/mtlasso/fold_5/", gene, ".GTEx_V8_fold_5_mtlasso.rds"))


    dat_input <- readRDS(paste0("data/cis_eqtl_analysis_ready/", gene, ".GTEx_V8.rds"))

    sample_size <- apply(dat_input$y_res, 2, function(x){sum(!is.na(x))})
    sample_size <- data.frame(tissue=names(sample_size), sample_size)

    ###enet accuracy
    acc_enet1 <- compute_accuracy(dat1_first$Ytest, dat1_first$Yhat_test_glmnet)
    r2_enet1 <- acc_enet1$r2
    scaled_rmse_enet1 <- acc_enet1$scaled_rmse

    acc_enet2 <- compute_accuracy(dat2_first$Ytest, dat2_first$Yhat_test_glmnet)
    r2_enet2 <- acc_enet2$r2
    scaled_rmse_enet2 <- acc_enet2$scaled_rmse

    acc_enet3 <- compute_accuracy(dat3_first$Ytest, dat3_first$Yhat_test_glmnet)
    r2_enet3 <- acc_enet3$r2
    scaled_rmse_enet3 <- acc_enet3$scaled_rmse

    acc_enet4 <- compute_accuracy(dat4_first$Ytest, dat4_first$Yhat_test_glmnet)
    r2_enet4 <- acc_enet4$r2
    scaled_rmse_enet4 <- acc_enet4$scaled_rmse

    acc_enet5 <- compute_accuracy(dat5_first$Ytest, dat5_first$Yhat_test_glmnet)
    r2_enet5 <- acc_enet5$r2
    scaled_rmse_enet5 <- acc_enet5$scaled_rmse

    r2_enet <- cbind(r2_enet1, r2_enet2, r2_enet3, r2_enet4, r2_enet5)
    scaled_rmse_enet <- cbind(scaled_rmse_enet1, scaled_rmse_enet2, scaled_rmse_enet3, scaled_rmse_enet4, scaled_rmse_enet5)

    ###mr.mash first pass accuracy
    acc_mrmash1 <- compute_accuracy(dat1_first$Ytest, dat1_first$Yhat_test)
    r2_mrmash1 <- acc_mrmash1$r2
    scaled_rmse_mrmash1 <- acc_mrmash1$scaled_rmse

    acc_mrmash2 <- compute_accuracy(dat2_first$Ytest, dat2_first$Yhat_test)
    r2_mrmash2 <- acc_mrmash2$r2
    scaled_rmse_mrmash2 <- acc_mrmash2$scaled_rmse

    acc_mrmash3 <- compute_accuracy(dat3_first$Ytest, dat3_first$Yhat_test)
    r2_mrmash3 <- acc_mrmash3$r2
    scaled_rmse_mrmash3 <- acc_mrmash3$scaled_rmse

    acc_mrmash4 <- compute_accuracy(dat4_first$Ytest, dat4_first$Yhat_test)
    r2_mrmash4 <- acc_mrmash4$r2
    scaled_rmse_mrmash4 <- acc_mrmash4$scaled_rmse

    acc_mrmash5 <- compute_accuracy(dat5_first$Ytest, dat5_first$Yhat_test)
    r2_mrmash5 <- acc_mrmash5$r2
    scaled_rmse_mrmash5 <- acc_mrmash5$scaled_rmse

    r2_mrmash_first <- cbind(r2_mrmash1, r2_mrmash2, r2_mrmash3, r2_mrmash4, r2_mrmash5)
    scaled_rmse_mrmash_first <- cbind(scaled_rmse_mrmash1, scaled_rmse_mrmash2, scaled_rmse_mrmash3, scaled_rmse_mrmash4, scaled_rmse_mrmash5)

    ###mr.mash second pass accuracy
    acc_mrmash1 <- compute_accuracy(dat1_sec$Ytest, dat1_sec$Yhat_test)
    r2_mrmash1 <- acc_mrmash1$r2
    scaled_rmse_mrmash1 <- acc_mrmash1$scaled_rmse

    acc_mrmash2 <- compute_accuracy(dat2_sec$Ytest, dat2_sec$Yhat_test)
    r2_mrmash2 <- acc_mrmash2$r2
    scaled_rmse_mrmash2 <- acc_mrmash2$scaled_rmse

    acc_mrmash3 <- compute_accuracy(dat3_sec$Ytest, dat3_sec$Yhat_test)
    r2_mrmash3 <- acc_mrmash3$r2
    scaled_rmse_mrmash3 <- acc_mrmash3$scaled_rmse

    acc_mrmash4 <- compute_accuracy(dat4_sec$Ytest, dat4_sec$Yhat_test)
    r2_mrmash4 <- acc_mrmash4$r2
    scaled_rmse_mrmash4 <- acc_mrmash4$scaled_rmse

    acc_mrmash5 <- compute_accuracy(dat5_sec$Ytest, dat5_sec$Yhat_test)
    r2_mrmash5 <- acc_mrmash5$r2
    scaled_rmse_mrmash5 <- acc_mrmash5$scaled_rmse

    r2_mrmash_sec <- cbind(r2_mrmash1, r2_mrmash2, r2_mrmash3, r2_mrmash4, r2_mrmash5)
    scaled_rmse_mrmash_sec <- cbind(scaled_rmse_mrmash1, scaled_rmse_mrmash2, scaled_rmse_mrmash3, scaled_rmse_mrmash4, scaled_rmse_mrmash5)

    ###mtlasso accuracy
    acc_mtlasso1 <- compute_accuracy(dat1_mtlasso$Ytest, dat1_mtlasso$Yhat_test)
    r2_mtlasso1 <- acc_mtlasso1$r2
    scaled_rmse_mtlasso1 <- acc_mtlasso1$scaled_rmse

    acc_mtlasso2 <- compute_accuracy(dat2_mtlasso$Ytest, dat2_mtlasso$Yhat_test)
    r2_mtlasso2 <- acc_mtlasso2$r2
    scaled_rmse_mtlasso2 <- acc_mtlasso2$scaled_rmse

    acc_mtlasso3 <- compute_accuracy(dat3_mtlasso$Ytest, dat3_mtlasso$Yhat_test)
    r2_mtlasso3 <- acc_mtlasso3$r2
    scaled_rmse_mtlasso3 <- acc_mtlasso3$scaled_rmse

    acc_mtlasso4 <- compute_accuracy(dat4_mtlasso$Ytest, dat4_mtlasso$Yhat_test)
    r2_mtlasso4 <- acc_mtlasso4$r2
    scaled_rmse_mtlasso4 <- acc_mtlasso4$scaled_rmse

    acc_mtlasso5 <- compute_accuracy(dat5_mtlasso$Ytest, dat5_mtlasso$Yhat_test)
    r2_mtlasso5 <- acc_mtlasso5$r2
    scaled_rmse_mtlasso5 <- acc_mtlasso5$scaled_rmse

    r2_mtlasso <- cbind(r2_mtlasso1, r2_mtlasso2, r2_mtlasso3, r2_mtlasso4, r2_mtlasso5)
    scaled_rmse_mtlasso <- cbind(scaled_rmse_mtlasso1, scaled_rmse_mtlasso2, scaled_rmse_mtlasso3, scaled_rmse_mtlasso4, scaled_rmse_mtlasso5)

    ###Combined accuracy

    mean_r2 <- cbind(enet=rowMeans(r2_enet, na.rm=TRUE), mrmash_first=rowMeans(r2_mrmash_first, na.rm=TRUE), 
                     mrmash_sec=rowMeans(r2_mrmash_sec, na.rm=TRUE), mtlasso=rowMeans(r2_mtlasso, na.rm=TRUE))
    se_r2 <- cbind(enet=matrixStats::rowSds(r2_enet, na.rm=TRUE)/apply(r2_enet, 1, function(x){sum(is.finite(x))}), 
                    mrmash_first=matrixStats::rowSds(r2_mrmash_first, na.rm=TRUE)/apply(r2_mrmash_first, 1, function(x){sum(is.finite(x))}),
                    mrmash_sec=matrixStats::rowSds(r2_mrmash_sec, na.rm=TRUE)/apply(r2_mrmash_sec, 1, function(x){sum(is.finite(x))}),
                    mtlasso=matrixStats::rowSds(r2_mtlasso, na.rm=TRUE)/apply(r2_mtlasso, 1, function(x){sum(is.finite(x))}))
                    
    mean_scaled_rmse <- cbind(enet=rowMeans(scaled_rmse_enet, na.rm=TRUE), mrmash_first=rowMeans(scaled_rmse_mrmash_first, na.rm=TRUE), 
                              mrmash_sec=rowMeans(scaled_rmse_mrmash_sec, na.rm=TRUE), mtlasso=rowMeans(scaled_rmse_mtlasso, na.rm=TRUE))
    se_scaled_rmse <- cbind(enet=matrixStats::rowSds(scaled_rmse_enet, na.rm=TRUE)/apply(scaled_rmse_enet, 1, function(x){sum(is.finite(x))}), 
                            mrmash_first=matrixStats::rowSds(scaled_rmse_mrmash_first, na.rm=TRUE)/apply(scaled_rmse_mrmash_first, 1, function(x){sum(is.finite(x))}),
                            mrmash_sec=matrixStats::rowSds(scaled_rmse_mrmash_sec, na.rm=TRUE)/apply(scaled_rmse_mrmash_sec, 1, function(x){sum(is.finite(x))}),
                            mtlasso=matrixStats::rowSds(scaled_rmse_mtlasso, na.rm=TRUE)/apply(scaled_rmse_mtlasso, 1, function(x){sum(is.finite(x))}))
    
    ###Plots
    min_lim <- min(c(mean_scaled_rmse[, "mrmash_first"], mean_scaled_rmse[, "mrmash_sec"]))
    max_lim <- max(c(mean_scaled_rmse[, "mrmash_first"], mean_scaled_rmse[, "mrmash_sec"]))
    p_mrmash_sec[[t]] <- ggplot(as.data.frame(mean_scaled_rmse), aes(x=mrmash_sec, y=mrmash_first)) + geom_point(shape=1) + xlim(min_lim, max_lim) + ylim(min_lim, max_lim) +
                                labs(x = "mr.mash (second)", y = "mr.mash (first)", title=gene) + geom_abline(intercept = 0, slope = 1) + theme_cowplot() +
                                theme(plot.title = element_text(hjust = 0.5))    

    min_lim <- min(c(mean_scaled_rmse[, "mrmash_first"], mean_scaled_rmse[, "enet"]))
    max_lim <- max(c(mean_scaled_rmse[, "mrmash_first"], mean_scaled_rmse[, "enet"]))
    p_enet[[t]] <- ggplot(as.data.frame(mean_scaled_rmse), aes(x=enet, y=mrmash_first)) + geom_point(shape=1) + xlim(min_lim, max_lim) + ylim(min_lim, max_lim) +
                    labs(x = "e-net", y = "mr.mash", title=gene) + geom_abline(intercept = 0, slope = 1) + theme_cowplot() +
                    theme(plot.title = element_text(hjust = 0.5))    

    min_lim <- min(c(mean_scaled_rmse[, "mrmash_first"], mean_scaled_rmse[, "mtlasso"]))
    max_lim <- max(c(mean_scaled_rmse[, "mrmash_first"], mean_scaled_rmse[, "mtlasso"]))
    p_mtlasso[[t]] <- ggplot(as.data.frame(mean_scaled_rmse), aes(x=mtlasso, y=mrmash_first)) + geom_point(shape=1) + xlim(min_lim, max_lim) + ylim(min_lim, max_lim) +
                        labs(x = "smt-lasso", y = "mr.mash", title=gene) + geom_abline(intercept = 0, slope = 1) + theme_cowplot() +
                        theme(plot.title = element_text(hjust = 0.5))    
}
```

### *mr.mash* first pass vs second pass

```{r mr.mash first vs second pass, message=FALSE, fig.height=15, fig.width=15}
plot_grid(p_mrmash_sec[[1]], p_mrmash_sec[[2]], p_mrmash_sec[[3]], p_mrmash_sec[[4]], p_mrmash_sec[[5]],
          p_mrmash_sec[[6]], p_mrmash_sec[[7]], p_mrmash_sec[[8]], p_mrmash_sec[[9]], NULL, 
          p_mrmash_sec[[10]], NULL,
          ncol = 3)
```

The results comparing the two versions of *mr.mash* show that the second pass does not improve the performance. Since it is also computationally demanding to perform the second pass, there is no reason to do it. 

### *mr.mash* first pass vs elastic net

```{r mr.mash first vs enet, message=FALSE, fig.height=15, fig.width=15}
plot_grid(p_enet[[1]], p_enet[[2]], p_enet[[3]], p_enet[[4]], p_enet[[5]],
          p_enet[[6]], p_enet[[7]], p_enet[[8]], p_enet[[9]], NULL,
          p_enet[[10]], NULL,
          ncol = 3)
```

The results comparing *mr.mash* and univariate elastic net show that, as expected, *mr.mash* achieves a lower RMSE for the majority of tissue/gene combinations. And even in the cases where the elastic net achieves a lower RMSE, *mr.mash*'s performance is very close.  

### *mr.mash* first pass vs sparse multi-task LASSO

```{r mr.mash first vs mtlasso, message=FALSE, fig.height=15, fig.width=15}
plot_grid(p_mtlasso[[1]], p_mtlasso[[2]], p_mtlasso[[3]], p_mtlasso[[4]], p_mtlasso[[5]],
          p_mtlasso[[6]], p_mtlasso[[7]], p_mtlasso[[8]], p_mtlasso[[9]], NULL,
          p_mtlasso[[10]], NULL,
          ncol = 3)
```
The results comparing *mr.mash* and sparse multi-task LASSO show that *mr.mash* performs better for five out of ten genes, and about the same in the remaining genes.

Overall, *mr.mash* seems to work well.
