---
title: "Convergence"
author: "Fabio Morgante"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

```{r set opts}
options(stringsAsFactors = FALSE)
```
#Simulation 1 -- Shared effects, independent variables

```{r load data 1}
dat1 <- readRDS("output/fit_mr_mash_n600_p1000_p_caus50_r5_pve0.5_sigmaoffdiag1_sigmascale0.8_gammaoffdiag0_gammascale0.8_Voffdiag0.2_Vscale0_updatew0TRUE_updatew0TRUE_updatew0methodmixsqp_updateVTRUE.rds")
n1 <- dat1$params$n
p1 <- dat1$params$p
p_causal1 <- dat1$params$p_causal
r1 <- dat1$params$r
pve1 <- dat1$params$pve
prop_testset1 <- dat1$params$prop_testset
progress_dat1 <- dat1$fit$progress
V1 <- dat1$inputs$V
Sigma1 <- dat1$inputs$Sigma
Gamma1 <- dat1$inputs$Gamma
```

The results below are based on simulation with `r n1` samples, `r p1` variables of which `r p_causal1` were causal, `r r1` responses with a per-response proportion of variance explained (PVE) of `r pve1`. Variables, X, were drawn from MVN(0, Gamma), causal effects, B, were drawn from MVN(0, Sigma). The responses, Y, were drawn from MN(XB, I, V).

```{r disp corrs 1}
cat("Gamma (First 5 elements)")
Gamma1[1:5, 1:5]

cat("Sigma")
Sigma1

cat("V")
V1
```

mr.mash was fitted to the training data (`r (1-prop_testset1)*100`% of the data) updating V and updating the prior weights using mixSQP.

Here, we investigate convergence.

```{r conv 1, fig.height=12, fig.width=15}
plot(progress_dat1$iter, progress_dat1$ELBO_diff, xlab="Iteration", ylab="log Difference in ELBO", main="ELBO vs iteration", type="b", pch=16, cex.lab=1.5, cex.axis=1.5, log="y")
```